---
title: "Descriptive statistics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{descriptive-statistics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette describes how to calculate basic descriptive statistics of the 1-dimensional data.

```{r setup}
suppressPackageStartupMessages(library(tidyverse))
library(bspm1d)

# import the 3 built-in datasets
data("simulated_two_local_max")
data("gait_symmetry")
data("plantar_arch_angle")
```

We will first demonstrate the functions with the dataset `simulated_two_local_max`. This dataset contains 12 simulated 1-dimensional curves (101 data points per curve). The first 6 rows are just white noise while the second 6 rows contain white noise plus two gaussian bursts at positions 25 and 75. Printing the object to the console reveals it is a nested tibble containing 3 columns: `ID` (subject identifier: 1-12), `group` (grouping factor: 1-2) and `Y` (tibbles of size 101 $\times$ 2).

```{r}
print(simulated_two_local_max)
```

The 1-dimensional data are nested inside the inner tibbles. Printing the first 10 rows shows they contain two columns: `time` (normalized to 0-100%) and `Y` (the outcome variable).

```{r}
print(simulated_two_local_max$Y[[1]])
```

How the user wishes to store the data (`matrix`, `data.frame`, (nested) `tibbles`, ...) is not important. What matters is that the functions in the package require the data be in long format. Below we demonstrate how to do that for nested tibbles.

```{r long-format}
data_long <- simulated_two_local_max %>%
  unnest(Y)

# print the first rows
print(data_long)
```

The dataset in long format now has 4 columns: `ID`, `group`, `time` and `Y`. This dataset can be used in the plotting functions. The basic plotting functions are `plot_raw1d()` and `plot_meanerr_1d()`. Figure \@ref(fig:plotmeansd) shows the basic functioning of the latter.

```{r plotmeansd, fig.cap="Descriptive statistics for the simulated dataset to illustrate the two-sample t-test.", out.width="100%", fig.asp=.75, fig.align="center"}
plot_meanerr_1d(data_long, outcome = "Y", 
                dimension = "time", grp_factors = "group") +
  labs(x = "Time (%)", y = "Y(t)", color = "Group", fill= "Group",
       title = "Simulated 1-dimensional data",
       caption = "Group 2 = smooth white noise\nGroup 1 = smooth white noise + gaussian bursts at t=25 and t=75.") +
  theme_minimal()
```





